---
title: "NYC CitiBike Trip Data"
author: "Wencong (Priscilla) Li (liwencong1995@gmail.com)"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to use NYC CitiBike Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE, message=FALSE}
library(knitr)
opts_chunk$set(fig.width = 6, fig.height = 4)
```

## NYC CitiBike Data 
The `citibike` package provides a interface for urser who are not experts of SQL to create and maintain an SQL database of [NYC CitiBike Data](https://www.citibikenyc.com/system-data). For `citibike` package ursers, a reliable data storage directory is all they need . 

Here are the variables included in NYC CitiBike dataset:
*Trip Duration (seconds)
*Start Time and Date
*Stop Time and Date
*Start Station Name
*End Station Name
*Station ID
*Station Lat/Long
*Bike ID
*User Type (Customer = 24-hour pass or 7-day pass user; Subscriber = Annual Member)
*Gender (Zero=unknown; 1=male; 2=female)
*Year of Birth

## Get started:
### Install packages
The `etl` package provides the generic framework for the `citibike` package. To install the `citibike` package, you must install etl first. Since the `citibike`  packages currently live on GitHub and not on CRAN, you have to install it manually.

```{r, eval=FALSE, message=FALSE}
install.packages("devtools", repos="http://cran.rstudio.com/")
devtools::install_github("beanumber/citibike")
```

### Load the package
This command loads both `etl` and `dplyr`.
```{r, message=FALSE}
library(citibike)
```

### Database
The `etl` function has three variables -- `object`, `dir`, and `db`.
For more information about `etl` function, please read the [etl doucumentation](https://github.com/beanumber/etl/blob/master/README.Rmd).

## Getting Started
The following code loads citibike data from 2015 January into a SQLite database.
```{r}
bikes <- etl("citibike")
bikes %>%
  etl_create(years = 2015, months = 1)
```

```{r}
trips <- tbl(bikes,"trips")
head(trips)

trips %>%
  group_by(year, origin) %>%
  summarise(N = n(), min.month = min(month), max.month = max(month)) %>%
  arrange(desc(N))
```

## More Examples



