---
title: "CitiBike"
author: "Wencong (Priscilla) Li (liwencong1995@gmail.com) & Weijia (Vega) Zhang (wzhang23@smith.edu)"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to use NYC CitiBike Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE, message=FALSE}
library(knitr)
opts_chunk$set(fig.width = 6, fig.height = 4)
```

## ETL

`citibike` is an R package that facilitates Extract - Transform - Load operations for [NYC CitiBike Data](https://www.citibikenyc.com/system-data). `citibike` inherits functions from [etl ](https://github.com/beanumber/etl/blob/master/README.Rmd), which is the parent package. Similarly to `etl`, `citibike` allows the user to pass the date arguments and get a populated SQL database back. 

## ETL citiBike

`citibike` is dependent on `etl`. To get started, both packages need to be loaded.

```{r, eval=FALSE, message=FALSE}
install.packages("devtools", repos="http://cran.rstudio.com/")
devtools::install_github("beanumber/citibike")
```

```{r, message=FALSE}
library(etl)
library(citibike)
```

Instantiate an `etl` object using a string **citibike** that determines the class of the object. Specifiy the local directory for data storage.
```{r}
bikes <- etl("citibike", dir = "local_path")
class(bikes)
```

## Connect to a local or remote database

`etl` works with a local or remote database, so as `citibike`. If no SQL source has been specified, a local `RSQLite` database will be created for the user. However, the user could also specify the path to an existing database using `dplyr::src_sql`.


## Extract

Extract takes `years` and `months` parameters and allows user to fetch the data of specific date of interest. Note that the data is updated monthly. The default date is the starting month, which is ** July 2013**. If the user enters any date before that, the user will get an error message notification. The user could check the raw folder in the local directory that's created before. 


```{r, eval= FALSE}
# default
etl_extract(bikes)

# a specific month
etl_extract(bikes, years = 2014, months = 9)

# duration 
etl_extract(bikes, years = 2015, moths = 3:6)

# invalid date will bring up error message
etl_extract(bikes, months = 3:5)
```



## Transform

By default, `etl_transform` takes July 2013 data file and transform the raw data into CSV files. Similar to `etl_extract`, the user could specify the dates. The user could check the load folder of the local directory that has been created.
```{r, eval = FALSE}
bikes %>%
  etl_transform()

```

## Load

Import the CSV files into SQL and populate the SQL database with the transformed data.

```{r, eval = FALSE}
bikes %>%
  etl_load()

```


## About the dataset 

All of the tables contain the following columns:

* Trip Duration (seconds)
* Start Time and Date
* Stop Time and Date
* Start Station Name
* End Station Name
* Station ID
* Station Lat/Long
* Bike ID
* User Type (Customer = 24-hour pass or 7-day pass user; Subscriber = Annual Member)
* Gender (0=unknown; 1=male; 2=female) and Year of Birth.

The user could choose which ones to look at for further analysis.

## Examples

```{r, message=FALSE}
library(lubridate)
library(dplyr)
```

The following code loads citibike data from 2015 January into a SQLite database.
```{r,  eval = FALSE}
bikes <- etl("citibike")
bikes %>%
  etl_create(years = 2015, months = 2)

trips <- tbl(bikes,"trips")
head(trips, 4)

trips <- trips %>%
  collect()

citibike_time <- trips %>%
  select(starttime, stoptime) %>%
  mutate(starttime = dmy_hm(starttime)) %>%
  mutate(stoptime = dmy_hm(stoptime)) %>%
  mutate(duration_time = as.numeric(stoptime-starttime))

citibike_time <- citibike_time %>%
  filter(duration_time >= 0)

mean(citibike_time$duration_time)

citibike_start <- trips %>%
  select(start.station.name, start.station.longitude, start.station.latitude) %>%
  rename(longitude = start.station.longitude) %>%
  rename(latitude = start.station.latitude)
  
if (require(leaflet)) {
  leaflet(data = citibike_start) %>% 
    addTiles() %>%
    addCircles()
}

citibike_end <- trips %>%
  select(end.station.name, end.station.longitude, end.station.latitude) %>%
  rename(longitude = end.station.longitude) %>%
  rename(latitude = end.station.latitude)

if (require(leaflet)) {
  leaflet(data = citibike_end) %>% 
    addTiles() %>%
    addCircles()
}
```


