---
title: "NYC CitiBike Trip Data"
author: "Wencong (Priscilla) Li (liwencong1995@gmail.com)"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to use NYC CitiBike Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE, message=FALSE}
library(knitr)
opts_chunk$set(fig.width = 6, fig.height = 4)
```

## NYC CitiBike Data 
The `citibike` package provides a interface for urser who are not experts of SQL to create and maintain an SQL database of [NYC CitiBike Data](https://www.citibikenyc.com/system-data). For `citibike` package ursers, a reliable data storage directory is all they need . 

Here are the variables included in NYC CitiBike dataset:

-Trip Duration (seconds)

-Start Time and Date

-Stop Time and Date

-Start Station Name

-End Station Name

-Station ID

-Station Lat/Long

-Bike ID

-User Type (Customer = 24-hour pass or 7-day pass user; Subscriber = Annual Member)

-Gender (0=unknown; 1=male; 2=female)

-Year of Birth

## Getting started:
### Install packages
The `etl` package provides the generic framework for the `citibike` package. To install the `citibike` package, you must install etl first. Since the `citibike`  packages currently live on GitHub and not on CRAN, you have to install it manually.

```{r, eval=FALSE, message=FALSE}
install.packages("devtools", repos="http://cran.rstudio.com/")
devtools::install_github("beanumber/citibike")
```

### Load the package
This command loads both `etl` and `dplyr`.
```{r, message=FALSE}
library(citibike)
library(lubridate)
library(dplyr)
```

### etl
The `etl` function has three variables -- `object`, `dir`, and `db`.
For more information about `etl` function, please read the [etl doucumentation](https://github.com/beanumber/etl/blob/master/README.Rmd).

### Populate a local database with citibike data
The following code loads citibike data from 2015 January into a SQLite database.
```{r}
bikes <- etl("citibike")
bikes %>%
  etl_create(years = 2015, months = 2)
```

```{r}
trips <- tbl(bikes,"trips")
head(trips, 4)

trips <- trips %>%
  collect()
```

```{r}
citibike_time <- trips %>%
  select(starttime, stoptime) %>%
  mutate(starttime = dmy_hm(starttime)) %>%
  mutate(stoptime = dmy_hm(stoptime)) %>%
  mutate(duration_time = as.numeric(stoptime-starttime))

citibike_time <- citibike_time %>%
  filter(duration_time >= 0)

mean(citibike_time$duration_time)
```

## Examples
```{r, eval = FALSE}
citibike_start <- trips %>%
  select(start.station.name, start.station.longitude, start.station.latitude) %>%
  rename(longitude = start.station.longitude) %>%
  rename(latitude = start.station.latitude)
  
if (require(leaflet)) {
  leaflet(data = citibike_start) %>% 
    addTiles() %>%
    addCircles()
}
```

```{r,  eval = FALSE}
citibike_end <- trips %>%
  select(end.station.name, end.station.longitude, end.station.latitude) %>%
  rename(longitude = end.station.longitude) %>%
  rename(latitude = end.station.latitude)

if (require(leaflet)) {
  leaflet(data = citibike_end) %>% 
    addTiles() %>%
    addCircles()
}
```


